---
title: "myProjectWeek4"
author: "Radhika Vikas Kulkarni"
date: "June 7, 2020"
output: html_document
---

## Final Project Summary:
    This project aims to foresee the way by which they did the exercise. The feature "classe" in the     
    training dataset identifies the same. I have used Random Forest classifier to classify the exercise  
    activities into 5 classes -A,B,C,D,E.
    
    I have used training and testing datasets from the source:  
    [Source](http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har). 
    
    I have followed following steps:
    
    * 1.	Loading the necessary R packages.
    * 2.	Downloading the training dataset from    
        [Training Data](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv)
    * 3.	Downloading the testing dataset from
         [Testing Data](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv)
    * 4.	Data cleaning and dimensionality reduction. 
        I removed the first 7 features as  1)X, 2)user_name, 3)raw_timestamp_part_1, 
        4)raw_timestamp_part_2, 5)cvtd_timestamp, 6)new_window, 7) num_window
    * 5.	I have partitioned training dataset into 80% Training and 20% Validation Data set.
    * 6.	I have used Random Forest classifier with 5-Fold cross validation to create my model.
    * 7.	I have used the trained model to predict the classes of validation data set and produced the 
        confusion matrix for the validation data set.
    * 8.	Finally I used the same model to predict the classes of given testing (Quiz) dataset. 


```{r}
# clear all the session parameters
rm(list=ls());
gc();

library("randomForest")

library("caret")


# Setting of the URL for the download of the training and testing datasets
# Source: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har

urlTrainData <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
urlTestData  <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

# Downloading the training and testing datasets
train <- read.csv(url(urlTrainData),na.strings = c(""," ","?","NA",NA,"#DIV/0!"))
test  <- read.csv(url(urlTestData),na.strings = c(""," ","?","NA",NA,"#DIV/0!"))


```
```{r}
# Cleaning data sets by removing columns with values=NA
train <- train[,colSums(is.na(train)) == 0]
test <- test[,colSums(is.na(test)) == 0]
dim(train)
dim(test)

# Dimentionality reduction- Removing first 7 features as 1)X, 2)user_name, 3)raw_timestamp_part_1, 4)raw_timestamp_part_2, 5)cvtd_timestamp, 6)new_window, 7) num_window 
trainRed <- train[,-c(1:7)]
testRed <- test[,-c(1:7)]
dim(trainRed)
dim(testRed)

# Partitioning of training data into 80% Training and 20% Validation Data set
trainPart <- createDataPartition(trainRed$classe, p = 0.8, list = FALSE)
trainFold <- trainRed[trainPart,]
validFold <- trainRed[-trainPart,]
dim(trainFold)
dim(validFold)

```
```{r}
# RandomForest Classification Model with 5-Fold Cross Validation
set.seed(200)
controlRF <- trainControl(method="cv", number=5, verboseIter=FALSE)
myModelRF <- train(classe ~ ., data=trainFold, method="rf",
                          trControl=controlRF)
myModelRF$finalModel

```

```{r}
# Validation Results
predValid <- predict(myModelRF, newdata=validFold)
confusionMatrix(predValid, validFold$classe)                 

```
```{r}
# Predicting the the Test(Quiz) data using the trained model of RandomForest- myModelRF
predTEST <- predict(myModelRF, newdata=testRed)
predTEST

```

